---
title: "Human data processing"
author: "Sebastián Sáenz"
date: "09/11/2020"
output: github_document

---

```{r libraries, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
```
## Dependencies
* `r R.version.string`
* `tidyverse` (v . `r packageVersion("tidyverse")`)


## 1. Human proteome

Quality control data was obtained from the `summary.txt` ouput form MetaLab.

```{r summary_df}
summary_df_human <- read.table("data_raw/human/summary.txt",
                 header = TRUE,
                 sep = "\t",
                 check.names = FALSE)
```

#### Number of peptide sequences identified

```{r peptides_plot,}
summary_df_human %>%
  filter(!`Peptide Sequences Identified` > 30000) %>%
ggplot(aes(x =`Peptide Sequences Identified`)) +
  geom_histogram(binwidth = 100) +
  scale_x_continuous(breaks = seq(0, 10000, by = 500)) +
  ylab("Number of samples") +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

summary_df_human %>%
  filter(!`Peptide Sequences Identified` > 30000) %>%
ggplot(aes(x =`Raw file`,
           y =`Peptide Sequences Identified`))+
  geom_point(stat = "identity") +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 1, angle = 90))
  
  
```

Total of peptides identified:
```{r}
sum(summary_df_human$`Peptide Sequences Identified`)
```



Summary:
```{r}
summary(summary_df_human$`Peptide Sequences Identified`)
```


#### Percentage of MS/MS Identified
```{r msms_plot}
ggplot(data = summary_df_human, aes(x =`MS/MS Identified [%]` )) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = seq(0,23, by = 1)) +
  scale_y_continuous(breaks = seq(0,70, by = 5)) +
  theme_classic()

summary_df_human %>%
ggplot(aes(x =`Raw file`,
           y =`MS/MS Identified [%]`))+
  geom_point(stat = "identity") +
  scale_y_continuous(expand = c(0,0), limits = c(0,25)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 1, angle = 90))
```

Summary:

```{r}
summary(summary_df_human$`MS/MS Identified [%]`)
```


#### Standard protein

We used two proteins as standard

* Enolase 1 [P00924](https://www.uniprot.org/uniprot/P00924)
* Alcohol dehydrogenase 1 [P00330](https://www.uniprot.org/uniprot/P00330)
```{r proteingroup}
pg_df_human <- read.table("data_raw/human/proteinGroups_clean.txt",
                 header = TRUE,
                 sep = "\t",
                 check.names = FALSE)

pg_df_human %>%
  filter(`Protein IDs` == "P00924" | `Protein IDs` == "P00330") %>%
  pivot_longer(cols = 2:281,
               names_to = "samples",
               values_to = "Int") %>%
  mutate(log_int = log10(Int)) %>%
  ggplot(aes(x = samples,
             y = log_int,
           colour =`Protein IDs`)) +
  geom_point() +
  theme_classic() +
  theme(axis.text.x = element_text(size = 1, angle = 90))

pg_df_human %>%
  filter(`Protein IDs` == "P00924" | `Protein IDs` == "P00330") %>%
  pivot_longer(cols = 2:281,
               names_to = "samples",
               values_to = "Int") %>%
  mutate(log_int = log10(Int)) %>%
  ggplot(aes(x = `Protein IDs`,
             y = log_int,
           colour =`Protein IDs`)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, color = "black") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 1, angle = 90))

```

* Enolase 1
```{r}
pg_df_human %>%
  filter(`Protein IDs` == "P00924") %>%
  pivot_longer(cols = 2:281,
               names_to = "samples",
               values_to = "Int") %>%
  summary()
```

* Alcohol dehydrogenase 1
```{r}
pg_df_human %>%
  filter(`Protein IDs` == "P00330") %>%
  pivot_longer(cols = 2:281,
               names_to = "samples",
               values_to = "Int") %>%
  summary()
```



## 2. Microbiome metaproteome

Quality control data was obtained from the `summary.txt` ouput form MetaLab.

```{r summary_df_micro}
summary_df_micro <- read.table("data_raw/microbiome/summary.txt",
                 header = TRUE,
                 sep = "\t",
                 check.names = FALSE)
```

#### Number of peptide sequences identified

```{r peptides_plot_micro,}
summary_df_micro %>%
  filter(!`Peptide Sequences Identified` > 30000) %>%
ggplot(aes(x =`Peptide Sequences Identified`)) +
  geom_histogram(binwidth = 100) +
  scale_x_continuous(breaks = seq(0, 10000, by = 500)) +
  ylab("Number of samples") +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

summary_df_micro %>%
  filter(!`Peptide Sequences Identified` > 30000) %>%
ggplot(aes(x =`Raw file`,
           y =`Peptide Sequences Identified`))+
  geom_point(stat = "identity") +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 1, angle = 90))
  
  
```

Total of peptides identified:
```{r}
sum(summary_df_micro$`Peptide Sequences Identified`)
```



Summary:
```{r}
summary(summary_df_micro$`Peptide Sequences Identified`)
```


#### Percentage of MS/MS Identified
```{r msms_plot_micro}
ggplot(data = summary_df_micro, aes(x =`MS/MS Identified [%]` )) +
  geom_histogram(binwidth = 0.5) +
  scale_x_continuous(breaks = seq(0,23, by = 1)) +
  scale_y_continuous(breaks = seq(0,70, by = 5)) +
  theme_classic()

summary_df_micro %>%
ggplot(aes(x =`Raw file`,
           y =`MS/MS Identified [%]`))+
  geom_point(stat = "identity") +
  scale_y_continuous(expand = c(0,0), limits = c(0,25)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 1, angle = 90))
```

Summary:

```{r}
summary(summary_df_micro$`MS/MS Identified [%]`)
```


#### Standard protein
```{r proteingroup_micro}
pg_df_micro <- read.table("data_raw/microbiome/proteinGroups_clean.txt",
                 header = TRUE,
                 sep = "\t",
                 check.names = FALSE)

pg_df_micro %>%
  filter(`Protein IDs` == "sp|P00924|;O2.UC17-1_GL0070718" | `Protein IDs` == "P00330") %>%
  pivot_longer(cols = 2:281,
               names_to = "samples",
               values_to = "Int") %>%
  mutate(log_int = log10(Int)) %>%
  ggplot(aes(x = samples,
             y = log_int,
           colour =`Protein IDs`)) +
  geom_point() +
  theme_classic() +
  theme(axis.text.x = element_text(size = 1, angle = 90))

pg_df_micro %>%
  filter(`Protein IDs` == "sp|P00924|;O2.UC17-1_GL0070718" | `Protein IDs` == "P00330") %>%
  pivot_longer(cols = 2:281,
               names_to = "samples",
               values_to = "Int") %>%
  mutate(log_int = log10(Int)) %>%
  ggplot(aes(x = `Protein IDs`,
             y = log_int,
           colour =`Protein IDs`)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, color = "black") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 1, angle = 90))

```

* Enolase 1 clustered with [Enolase (Slackia piriformis YIT 12062)](https://www.uniprot.org/blast/uniprot/B20201111A94466D2655679D1FD8953E075198DA80108F2H)

```{r}
pg_df_micro %>%
  filter(`Protein IDs` == "sp|P00924|;O2.UC17-1_GL0070718") %>%
  pivot_longer(cols = 2:281,
               names_to = "samples",
               values_to = "Int") %>%
  summary()
```

## My computer

```{r sessioninfo, message=FALSE}
sessionInfo()
```

